Depression and suicidality as credible signals of need

### 2021-06-08 Paper.rmd
Deleted much text that was commented out. Retained text on social responses to depression

<!-- Almost all individuals will suffer adversity, such as injury or loss of material or social resources, at some point in their lives. Psychological pain, such as sadness and low mood, probably evolved to motivate victims of adversity to shift their attention to the causes of the problem so as to mitigate its negative fitness consequences and to learn to avoid future such adverse events [@thornhill1989; @andrews2010; @Nesse1990]. In many types of adversity, social partners could help minimize the costs. Therefore signals of psychological pain, such as sad expressions and crying, probably evolved to indicate need [e.g., @balsters2013]. CurrentDraft-->

<!-- Although the relationship between negative events and depression is bidirectional (cite), negative events still predict depression when looking at events outside of one’s control, indicating that the connection is unlikely to be driven solely by individuals who are already depressed selecting into situations where negative events are likely to be common (Hammen, 2005; Kendler, Karkowski, & Prescott, 1999). -->

@@

@@

<!-- Other notable risk factors, like loss of a loved one or severe or prolonged illness, might seem less related to conflict. In these situations, however, individuals will often need increased investment from social partners, who might not be able, or want, to provide more investment than they already are. The ability for others to provide meaningful support to limit the harm in times of need [@SugiyamaSugiyama2003] combined with the ubiquity of conflicts of interests mean that many types of problems will eventually involve social conflict even if they did not start that way [@Hagen2003]. There is indeed increasing evidence that loss of a loved one is often followed by increased family conflict [see @SymeHagen2021 for a brief review]. -->

@@

<!-- Individuals who have suffered a negative life event will often need help to mitigate the fitness cost. To obtain this help, they must signal need, e.g., by verbal requests, sad facial expressions, and crying. But if they have conflicts of interest with social partners who could provide that help, and if their need is private information, their signals might not be believed if the signals, like crying, could easily be faked [@Zahavi1975; @mokkonen2016] and if the cost of believing a deceptive signal was high [@dawkins1978; @johnstone1993; @Hasson1994]. -->

<!-- Conflicts of interest are ubiquitous, even among closely related individuals. In species that rely heavily on cooperation, like our own, conflicts can arise over the levels of investment in a cooperative endeavor and the division of the resulting benefits. -->

@@

<!-- Formal models of bargaining highlight the role of private information regarding the value each party places the cooperative endeavor. Withholding or risking one's contributions to a cooperative endeavor can be conceptualized as credibly signaling private information about one's need. Cooperation generates benefits over time. Thus, withholding or risking contributions incurs a *cost of delay* -- the more one values the endeavor, the greater the cost of delay.  -->

@@

<!-- @Rosenthal1993 proposed a credible signaling model of what he, like Firth, viewed as 'gambling-type' suicide attempts. In the Rosenthal model, as long as that risk is greater than or equal to the net benefit a non-needy person would realize by sending the signal (discounted by the risk of death), social partners can trust that signal senders are genuinely in need despite conflicts of interest [@Rosenthal1993].  -->

@@

<!-- As they are prohibitively costly for those in minor need, the costs of depression and suicidal behavior may be particularly important in times of substantial conflicts of interest or when the level of support needed to address one's situation is substantial, both of which should be expected to increase the payoff of manipulating others through fake signals. When either condition applies, less costly signals like verbal requests, facial sadness, and crying are likely to be less reliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals so long as the costs of signaling are outweighed by the benefits [@SymeHagen2021]. -->

<!-- Alternative start to line 92: Although verbal requests, sad facial expressions, and crying can be used to signal need, they are likely to be less effective in times of severe conflicts of interests or times when the level of support needed to address one's situation is substantial. This is because both conditions increase the payoff of manipulating others through fake signal.  -->

<!-- Both depression and suicide attempts are more common outcomes of these sorts of conflict than is often appreciated by the biomedical approach (Syme and Hagen 2021). With notable exceptions (Kendler; Mazure; Hammen), the biomedical model tends to frame both suicide and depression as a problem with the individual, e.g., as one involving psychiatric disturbance, drug addiction, or stress sensitivity [@wray2011 (changed wording to include depression]. This theoretical approach has made little progress: rates of suicide, which vary widely by country, have remained relatively flat in countries for which there are long-term data, such as the US, UK, Japan, and Sweden [@owidsuicide] (tweak for depression). -->

<!-- ## Depression and suicide attempts as both signals and bargaining tools -->

@@

<!-- Another possible explanation for the connection between depression, suicide, and social adversity is that some instances of depression and suicide attempts serve as means to elicit increased support from others by honestly signaling need to those invested in the well-being of the signaler, [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], something which may be particularly useful in times when an individual cannot solve substantial fitness-relevant challenges on their own [@Hagen2003; @SymeEtAl2016]. -->

<!-- Due to the ease with which signals can be faked [@Zahavi1975; @mokkonen2016] and the costs of accepting a deceptive signal as reliable [@dawkins1978; @johnstone1993; @Hasson1994], demonstrating that one's need is genuine has likely been a recurrent adaptive problem since humans evolved the capacity to provide instrumental support to one another. One potential solution to this is to produce signals that are unprofitable to fake [@Zahavi1975]. This is possible whenever the benefits of signaling outweigh the costs for those who have the relevant underlying quality but the costs outweigh the benefits for those who lack the quality or possess it to a lesser degree [@zahavi1977; @grafen1990]. In particular, if the benefits of sending the signal are the same for signalers regardless of condition, but the signal cost outweighs this benefit for signalers without the underlying quality (a "costly signal"), then individuals with the underlying quality can evolve to send the signal and receivers can evolve to respond in mutually beneficial ways to the signal. -->

<!-- Depression and suicidal behavior fit this criterion as both are relatively costly for those not facing adversity than those who are. In the case of depression, the large-scale reduction in one's normal routine of fitness enhancing behavior is relatively less costly for those whose situations result in a limited ability to obtain fitness-relevant benefits, with depression likely adding no additional costs in instances of extreme deprivation [@Hagen2003]. When it comes to suicide, the bulk of the costs associated with suicide attempts are only felt in times that result in death or severe bodily harm. For this reason, the way others view and respond to suicide attempts is likely to be strongly influenced by the probability that the attempt would result in death or severe injury [@Rosenthal1993]. Despite this difference, the previously mentioned logic still applies with increased risk of death being relatively more costly for those able to advance their fitness prospects in their day-to-day lives. -->

<!-- The costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one's situation is substantial, both of which should be expected to increase the payoff of manipulating others through fake signals. When either condition applies, less costly signals like verbal requests, facial sadness, and crying are likely to be less reliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals so long as the costs of signaling are outweighed by the benefits [@SymeHagen2021]. -->

<!-- ## Conflict plays a central role in suicidality and depression -->

<!-- Early on, anthropologists also tackled the problem of suicide as manifested in the small traditional societies in which they conducted their fieldwork. Suicide, they found, was commonly a form of protest, revenge, and/or appeal [e.g., @Malinowski:1932; @niehaus2012]. Some ethnographers, such as @Hezel:1987, conceptualized suicide as a form of anger, whereas others, such as @Giddens:1964, conceptualized suicide as a form of social pressure. @Counts:1980 emphasized the powerlessness of suicide victims. Common to almost all theoretical and empirical investigations of suicide is a focus on *completed* suicides, i.e., suicide deaths. @SymeEtAl2016, in contrast, argued that because the vast majority of suicidal behavior does not result in death, the proper focus should be on suicide ideation and suicide attempts (see Figure @ref(fig:cdc)). Suicide attempts, as we explain next, are a gamble to improve one's life by taking a low but non-zero risk of death. Most suicides deaths, especially in young, physically healthy individuals, are therefore the inevitable consequence that some individuals lose this gamble. -->

<!-- Within anthropology, Raymond Firth was likely the first to conceptualize suicidality as a gamble to improve one's circumstances in the here and now. Based on observations that suicide attempts often followed social conflict or loss, and varied substantially in their likelihood of death, he argued that a sizable subset of the suicide attempts among the Tikopia were not meant to end in death but instead were a means to elicit attention, status, or immediate reintegration into the community following negative events [@firth1936; @Firth:1961]. Human communites are highly interdependent [\@] -->

<!-- Suicidality is closely linked to depression. Suicidal ideation is one of nine diagnostic criteria for a major depressive episode [MDE, @dsm5], it is associated with depression across cultures [@haroz2017], and depression is a major risk factor for suicidal behavior [@hawton2013]. Compared to suicide, however, there are fewer anthropological investigations of depression [@kleinman1985] despite frequent reports of depression or depression-like behavior in a wide range of societies [@haroz2017; @liu2020] and the potential for cross-cultural research to help understand depressive states [@haroz2017]. -->

<!-- and anthropological explanations for low mood have become more common. In general, these approaches see sadness and minor depression as ways to motivate adaptive behavior changes in response to negative outcomes in one's environment (cite Nesse and others), mainly differing in how they view major depression. While some approaches present most forms of major depression as dysfunctional (inflammation and some of Nesse), others have extended the logic on the benefits of low mood and reduced behavior to more severe instances of sadness. Of the possible benefits, the ability of depression to help solve fitness-relevant in times of social conflict is the most emphasized, with depression being suggested to help by reallocating resources with the body for problem solving (cite ARH), reducing the costs of interacting with higher status individuals (cite), or serving as a means to both signal one's need and bargain for increased support (cite).  -->

@@

<!-- and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to  the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others). -->

<!-- Due to depression's relationship to suicide [@who2001; @hawton2013], it should not be surprising that suicidality also often coincides with social adversity and conflict. Grief over the loss of a loved one [@firth1936: @Firth:1961; @devries2011], bullying [@alavi2017; @butler2019; @ford2017; @wang2020], and instances of social transgression [@Firth:1961; @Malinowski:1932; @SymeHagen2018; @lester1997] have all been reported to lead to suicidal behavior in both industrialized and traditional societies. Of the recorded instances of suicide attempts in the Human Relations Area Files (HRAF), 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone [@SymeEtAl2016]. As with depression, assault is an particularly powerful risk factor for suicidal behavior based on multiple large-scale multi-country studies [@devries2011; @ellsberg2008] and more localized examinations [@reigstad2017]. -->

<!-- ## Depression and suicide attempts as both signals and bargaining tools -->

<!-- One possible reason for the connection between depression, suicide, and social adversity is that some instances of depression and suicide attempts serve as means to elicit increased support from others [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], particularly in times when an individual cannot solve substantial fitness-relevant challenges on their own [@Hagen2003; @SymeEtAl2016]. According to the bargaining models of depression and suicide, the small groups humans lived in for much of our evolutionary history created many situations in which certain forms of cooperation were unlikely to be easily replaced by others while also allowing for harmful or unfair social relationships to develop [@Hagen2003; @SymeEtAl2016]. In these instances, one's monopoly on one or more types of cooperation could be leveraged to change how one is treated if others valued an individual's cooperation sufficiently and believed them to be in genuine need. -->

<!-- Due to the ease with which signals can be faked [@Zahavi1975; @mokkonen2016] and the costs of accepting a deceptive signal as reliable [@dawkins1978; @johnstone1993; @Hasson1994], demonstrating that one's need is genuine has likely been a recurrent adaptive problem since humans evolved the capacity to provide instrumental support to one another. One potential solution to this is to produce signals that are unprofitable to fake [@Zahavi1975]. This is possible whenever the benefits of signaling outweigh the costs for those who have the relevant underlying quality but the costs outweigh the benefits for those who lack the quality or possess it to a lesser degree [@zahavi1977; @grafen1990]. In particular, if the benefits of sending the signal are the same for signalers regardless of condition, but the signal cost outweighs this benefit for signalers without the underlying quality (a "costly signal"), then individuals with the underlying quality can evolve to send the signal and receivers can evolve to respond in mutually beneficial ways to the signal. -->

<!-- Depression and suicidal behavior fit this criterion as both are relatively costly for those not facing adversity than those who are. In the case of depression, the large-scale reduction in one's normal routine of fitness enhancing behavior is relatively less costly for those whose situations result in a limited ability to obtain fitness-relevant benefits, with depression likely adding no additional costs in instances of extreme deprivation [@Hagen2003]. When it comes to suicide, the bulk of the costs associated with suicide attempts are only felt in times that result in death or severe bodily harm. For this reason, the way others view and respond to suicide attempts is likely to be strongly influenced by the probability that the attempt would result in death or severe injury [@Rosenthal1993]. Despite this difference, the previously mentioned logic still applies with increased risk of death being relatively more costly for those able to advance their fitness prospects in their day-to-day lives. -->

<!-- The costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one's situation is substantial, both of which should be expected to increase the payoff of manipulating others through fake signals. When either condition applies, less costly signals like verbal requests, facial sadness, and crying are likely to be less reliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals so long as the costs of signaling are outweighed by the benefits [@SymeHagen2021]. -->

<!-- Beyond signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support [@Hagen2003]. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from suicide. In these instances, depression and suicide attempts may function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, they would need to increase support towards the signaler. Due to this bargaining component, aversive responses from the targets of the signal to the signaler are expected throughout the bargaining process, with potential helpers likely signaling the costliness of providing support. Importantly, such responses do not prevent the delivery of meaningful benefits, they are simply expected to reduce the risk of helpers providing too much support for too little need. -->

<!-- ### Determining the impact of cues versus signals -->

<!-- Distinguishing between the information provided by signals versus cues is a major challenge for signaling approaches to behavior. A signal differs from a cue in that it is intentionally sent by an organism to alter the perceptions and behaviors of others, while a cue is any piece of information that influences an individual's behavior or physiological state that was not created for communicative effect [@MaynardSmithHarper1995]. Regarding the signals of interest, suicide attempts are more likely to take the form of pure, external signals than other symptoms of depression. This is because low mood's impact on behavior for non-signaling benefits likely conveys information as cues in addition to what might exist for signaling purposes. In other words, the signaling component of depression may come from the upregulation of symptoms beyond what would be optimal for non-signaling purposes rather than the creation of entirely new symptoms. Suicidality is likely an exception to this, due to the likelihood most suicide attempts do not have non-signaling functions [see @deCatanzaro1991 for a possible exception]. -->

<!-- ## Responses to depression and suicidal behavior  -->

<!-- For depression and suicide attempts to be worth the costs involved with signal creation, meaningful social support would be required [@Hagen2003; @SymeEtAl2016]. Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning [@EvraireDozois2011; @GotlibLee1989; @HirschfeldEtAl2000]. One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others [@JoinerMetalsky1995; @StarrDavila2008]. Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place [@MesserGross1995; @SchwartzEtAl2012]. -->

<!-- Despite being looked for less frequently, beneficial responses to depressed individuals are also reported. Depression has been associated with reduced likelihood of aggression within families [@hops1987; @dadds1992], including those already experiencing high levels of conflict [@SheeberEtAl2001]. There is also limited evidence that depression can result in increased levels of support from others outside the family. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together [@HokansonEtAl1986]. Likewise, a study examining responses to confederates trained to appear ‘depressed’, ‘anxious’, or ‘normal’ found that individuals were more likely to provide advice and support to ‘depressed’ individuals than ‘anxious’ or ‘normal’ individuals despite the confederate in the ‘depression’ condition being viewed more negatively than the other conditions and the ‘depression’ condition involving similar levels of distress compared to the ‘anxious’ condition [@stephens1987].  -->

<!-- Most studies examining responses to those who have survived suicide attempts have focused on stigmatization. In these studies, perceptions of survivors as being weak, selfish, mentally ill, and antisocial are all being commonly reported [@BatterhamEtAl2013; @TzengLipson2004], with stigmatization often being found within and outside one's social network [@FreyEtAl2016; @ScoccoEtAl2012]. Despite this potential for stigmatization, increased social support and changes to important relationships have been reported to follow suicide attempts with some indication that these effects may hold long term [@Stengel1956]. For example, a study of 100 women who survived suicide attempts found that individuals gained identifiable benefits through the attempt in 75 cases, with 41 individuals benefiting from reconciliations with others [@Lukianowicz1971]. Unlike many Western countries, where suicide is often viewed as pathological [@Hidaka2012], members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness [@Shostak1981], with attempts having been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers [@GutierrezdePinedaMuirden1948; @Hilger1957; @Karsten1935; @Tessmann1930; @Wilson1960]. Although findings that individuals view suicide attempts as cries for help is not necessarily evidence for their ability to lead to beneficial responses, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, despite suicide rarely if ever being the main topic [@SymeEtAl2016]. -->

@@

<!-- Overall, our results on the connection between belief in need and likelihood of help are consistent with costly signaling and bargaining approaches to depression and suicide. At T1, *Belief* and *Action* were highly correlated yet distinct; while at T2, the effect of signal on both *Belief* and *Action* was to increase the magnitude of pro-signaler responses, with the effect of signal on *T2 Action* being largely mediated by signal's effect on belief in need. This relationship between belief in need and likelihood of support and the importance of signal cost should be expected in scenarios when a target of a signal of need would benefit from helping if the signaler's need was genuine but face substantial net costs if it was not. Importantly, such scenarios appear to be common among modern humans and have likely been since individuals could increase their inclusive fitness through supporting kin or important cooperative partners who could return to providing benefits to supporters once their situation returned to normal [@SugiyamaSugiyama2003; @tooby1996]. -->

<!-- Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are consistent with this hypothesis. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception and the number of defectors within the broader population [@Rosenthal1993], with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact that depression and suicide attempts become relatively less costly to one's fitness prospects with increased need [@Hagen2003], suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose. -->

@@

<!-- Slider anchor  -->
<!-- Regex: stopped in SI -->
<!-- T\d \w+ -->

@@

 <!-- The need to explain suicide's connection to depression is often overlooked by explanations for major depression, particularly those which propose non-signaling functions. One explanation for their connection, is that both suicidality and depression may function to limit negative fitness outcomes in times of substantial adversity [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], with costlier signals being required to elicit support when the payoffs to faking need are high enough so that less costly signals are not reliable indicators of one's need [@Rosenthal1993]. For this hypothesis to be correct, instrumental support would need to be provided frequently enough with large enough impacts so that the costs of signaling are outweighed by the benefits on average. In support of this, we found exposure to depression and suicide attempts resulted in greater belief in need and increased likelihood of support in fictional scenarios. -->

@@

<!-- # To do: -->

<!-- -   victim or signaler? -->
<!-- -   add anger -->
<!-- -   Intro "State the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results." -->
<!-- -   Discussion " This should explore the significance of the results of the work, not repeat them. A combined Results and Discussion section is often appropriate. Avoid extensive citations and discussion of published literature." -->
<!-- -   mention asking for prefered sex -->

<!-- # Notes from my meeting with Anne -->

<!-- -   Table detailing the cooperative endeavor (maybe word as boundedness of fate), conflict of interest, and what the private information is about would be useful. -->
<!-- -   Calling verbal request control when it's a treatment? -->
<!-- -   Can shift much of the pilot studies info to SI with some detail post line 279 in draft she saw. -->
<!-- -   Cumulative distribution plot: needs more info in captions. could benefit from fig6a and fig6b (confusion over x and y axes) -->

>>>>

@@

## Responses to depression and suicidal behavior

For depression and suicide attempts to be worth the costs involved with signal creation, meaningful social support would be required [@Hagen2003; @SymeEtAl2016]. Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning [@EvraireDozois2011; @GotlibLee1989; @HirschfeldEtAl2000]. One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others [@JoinerMetalsky1995; @StarrDavila2008]. Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place [@MesserGross1995; @SchwartzEtAl2012].

Despite being looked for less frequently, beneficial responses to depressed individuals are also reported. Depression has been associated with reduced likelihood of aggression within families [@hops1987; @dadds1992], including those already experiencing high levels of conflict [@SheeberEtAl2001]. There is also limited evidence that depression can result in increased levels of support from others outside the family. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together [@HokansonEtAl1986]. Likewise, a study examining responses to confederates trained to appear ‘depressed’, ‘anxious’, or ‘normal’ found that individuals were more likely to provide advice and support to ‘depressed’ individuals than ‘anxious’ or ‘normal’ individuals despite the confederate in the ‘depression’ condition being viewed more negatively than the other conditions and the ‘depression’ condition involving similar levels of distress compared to the ‘anxious’ condition [@stephens1987].

Most studies examining responses to those who have survived suicide attempts have focused on stigmatization. In these studies, perceptions of survivors as being weak, selfish, mentally ill, and antisocial are all being commonly reported [@BatterhamEtAl2013; @TzengLipson2004], with stigmatization often being found within and outside one's social network [@FreyEtAl2016; @ScoccoEtAl2012]. Despite this potential for stigmatization, increased social support and changes to important relationships have been reported to follow suicide attempts with some indication that these effects may hold long term [@Stengel1956]. For example, a study of 100 women who survived suicide attempts found that individuals gained identifiable benefits through the attempt in 75 cases, with 41 individuals benefiting from reconciliations with others [@Lukianowicz1971]. Unlike many Western countries, where suicide is often viewed as pathological [@Hidaka2012], members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness [@Shostak1981], with attempts having been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers [@GutierrezdePinedaMuirden1948; @Hilger1957; @Karsten1935; @Tessmann1930; @Wilson1960]. Although findings that individuals view suicide attempts as cries for help is not necessarily evidence for their ability to lead to beneficial responses, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, despite suicide rarely if ever being the main topic [@SymeEtAl2016].

## Depression, anger, and conflict

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@
